---
title: "Testing the New Interface with the OU Model"
author: "Martin Lysy"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: yes
  rmarkdown::html_document:
    toc: yes
    theme: journal
vignette: >
  %\VignetteIndexEntry{Testing the New Interface with the OU Model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

\newcommand{\ud}{\mathop{}\!\mathrm{d}}
\newcommand{\cov}{\operatorname{cov}}
\newcommand{\psd}{\mathcal{S}}
\newcommand{\f}{f}

# Model

The OU model is given by the stochastic differential equation
$$
\ud X_t = -\alpha X_t \ud t + \beta \ud B_t.
$$
The autocorrelation function is $\cov(X_s, X_{s+t}) = \beta^2/(2\alpha) \cdot e^{-\alpha|t|}$, and the PSD is
$$
\psd_X(\f) = \frac{\beta^2}{(2\pi \f)^2 + \alpha^2}.
$$

# Simulate Data

```{r setup}
require(realPSD)
require(TMB)
require(R6)
require(tidyverse)
```

```{r sim}

#' PSD for the OU model.
#'
#' @param freq Vector of frequencies
#' @param alpha Mean reversion parameter.
#' @param beta Scale parameter.
#' @return PSD vector at the frequencies in `freq`.
ou_psd <- function(freq, alpha, beta) {
  beta^2/((2*pi*freq)^2 + alpha^2)
}

# true parameter values
theta0 <- c(alpha = 3.1, beta = 2.2)

# timescale for simulation
frange <- c(1e-4, 1e2) # frequency range
fs <- frange[2]*2 # sampling frequency
dt <- 1/fs # interobservation time
N <- floor(fs/frange[1]) # number of observations

# time domain simulation
Xt <- ou_sim(gamma = theta0[1], mu = 0, sigma = theta0[2], dt = dt, n_obs = N)
```

## Empirical vs Theoretical PSD Estimate

```{r psd_emp}
psd_emp <- periodogram(Xt, fs = fs)
psd_emp <- psd_emp %>% filter(freq < 20)
bin_size <- 100 # for estimation and plotting
psd_bin <- as.data.frame(lapply(psd_emp, binning,
                                bin_size = bin_size, bin_type = "mean"))

plot(psd_bin$freq, psd_bin$Ypsd, type = "l", log = "xy",
     xlab = "Frequency (Hz)", ylab = "PSD")
curve(ou_psd(freq = x, alpha = theta0[1], beta = theta0[2]),
      col = "red", add = TRUE)
```

# Create the `ou_model` Object

The first step is to create a standalone version of the model.  Here is the definition of the `ou::UFun` class:

```{r ou_ufun, echo = FALSE, results = "asis"}
cat("```cpp", readLines("OU_Model.hpp"), "```", sep = "\n")
```

Now let's compile it and create the `ou_model` object as follows:

```{r ou_cpp_def}
# create c++ file
ou_cpp <- make_psd_model(model = "OU",
                         header = "OU_Model.hpp",
                         class = "ou::UFun",
                         ctor = "ou::make_Ufun",
                         standalone = TRUE)
# this avoid compiling if file hasn't changed
if(!file.exists("OU_FitMethods.cpp") ||
   !identical(ou_cpp, paste0(readLines("OU_FitMethods.cpp"), collapse = "\n"))) {
  cat(ou_cpp, sep = "\n", file = "OU_FitMethods.cpp")
}

# compile it
model <- "OU_FitMethods"
TMB::compile(paste0(model, ".cpp"),
             PKG_CXXFLAGS = paste0("-I", system.file("include", package = "realPSD"),
                                   " -std=c++11"))
dyn.load(TMB::dynlib(model))

# create the ou_model class
ou_model <- R6::R6Class(
  classname = "ou_model",
  inherit = realPSD::psd_model,
  private = list(
    n_phi_ = 1,
    tmb_model_ = "OU",
    tmb_DLL_ = "OU_FitMethods"
  )
)
```

## Testing

To check whether the implementation is correct, we'll compare the **TMB** implementation of the $U(\f, \phi)$ method of `ou_model` to a pure R functional implementation.

```{r tmb_test}
# construct the model object
ou_obj <- ou_model$new(freq = psd_emp$freq, Ypsd = psd_emp$Ypsd,
                       bin_size = 100, est_type = "lp")

ou_ufun <- ou_obj$ufun() # TMB method corresponding to UFun.eval()
U_tmb <- ou_ufun$fn(log(theta0[1])) # evaluate at frequencies for lp estimator
# same calculation in R
U_r <- ou_psd(freq = psd_bin$freq, alpha = theta0[1], beta = 1)
range(U_tmb - U_r)
```

# Parameter Estimation

## MLE Method

```{r mle_fit}

#` Helper function to convert a value of `phi_est` into `eta_coef` and `eta_vcov`.
#` @param phi Value of `phi`.
#` @param obj Model object created by `ou_model`.
#` @return List with elements `coef` and `vcov` of parameters on the computational basis, `eta = (phi, zeta)`.
to_fit <- function(phi, obj) {
  fit <- list(phi = phi, zeta = obj$nlp()$simulate(phi)$zeta)
  fit <- list(coef = c(phi = fit$phi, zeta = fit$zeta),
              vcov = obj$vcov(phi = fit$phi, zeta = fit$zeta))
  rownames(fit$vcov) <- colnames(fit$vcov) <- names(fit$coef)
  fit
}

# instantiate TMB object for profile likelihood
ou_obj$set_est(est_type = "mle")
ou_nlp <- ou_obj$nlp()

opt <- optimize(f = ou_nlp$fn, lower = log(.1), upper = log(10))
mle_fit <- to_fit(opt$min, ou_obj)

plot(psd_bin$freq, psd_bin$Ypsd, type = "l", log = "xy",
     xlab = "Frequency (Hz)", ylab = "PSD")
curve(ou_psd(freq = x, alpha = exp(mle_fit$coef[1]), beta = exp(mle_fit$coef[2]/2)),
      col = "red", add = TRUE)
```

## LP Method

```{r lp_fit}
# instantiate TMB object for profile likelihood
ou_obj$set_est(est_type = "lp", bin_size = bin_size)
ou_nlp <- ou_obj$nlp()

opt <- optimize(f = ou_nlp$fn, lower = log(.1), upper = log(10))
lp_fit <- to_fit(opt$min, ou_obj)

plot(psd_bin$freq, psd_bin$Ypsd, type = "l", log = "xy",
     xlab = "Frequency (Hz)", ylab = "PSD")
curve(ou_psd(freq = x, alpha = exp(lp_fit$coef[1]), beta = exp(lp_fit$coef[2]/2)),
      col = "red", add = TRUE)
```

## NLS Method

```{r nls_fit}
# instantiate TMB object for profile likelihood
ou_obj$set_est(est_type = "nls", bin_size = bin_size)
ou_nlp <- ou_obj$nlp()

opt <- optimize(f = ou_nlp$fn, lower = log(.1), upper = log(10))
nls_fit <- to_fit(opt$min, ou_obj)

plot(psd_bin$freq, psd_bin$Ypsd, type = "l", log = "xy",
     xlab = "Frequency (Hz)", ylab = "PSD")
curve(ou_psd(freq = x, alpha = exp(lp_fit$coef[1]), beta = exp(lp_fit$coef[2]/2)),
      col = "red", add = TRUE)
```

## Table of Parameter Estimates

```{r disp}
disp <- sapply(list(mle = mle_fit, lp = lp_fit, nls = nls_fit),
               function(fit) c(fit$coef[1], sqrt(fit$vcov[1,1]),
                               fit$coef[2], sqrt(fit$vcov[2,2])))
disp <- cbind(true = c(log(theta0[1]), NA, 2 * log(theta0[2]), NA),
              disp)
disp <- t(disp)
colnames(disp) <- c("phi_est", "phi_se", "zeta_est", "zeta_se")
signif(disp, 3)
```

# Scratch

The code below was executed during the debugging of this file.  It is not run here, only kept for reference.

```{r scratch, eval = FALSE}

curve(sapply(x, function(phi) ou_nlp$fn(phi)), from = log(.1), to = log(10))

# ok check whether we get the same thing with pure R
source(system.file("tests", "testthat", "realPSD-testfunctions.R",
                   package = "realPSD"))

ou_ufun_r <- function(f, phi) {
  alpha <- exp(phi)
  1/((2*pi*f)^2 + alpha^2)
}
ou_nlp_r <- function(phi) {
  mle_nlp_r(phi, Y = psd_bin$Ypsd, f = psd_bin$freq,
            ufun = ou_ufun_r, fs = 1)
}
ou_nll_r <- function(eta) {
  lp_nll_r(phi = eta[1], zeta = eta[2], Zbar = log(psd_bin$Ypsd), fbar = psd_bin$freq,
           ufun = ou_ufun_r, fs = 1, B = bin_size, bin_type = "mean")
}

ou_nlp_r(log(theta0[1]))
ou_nlp$fn(log(theta0[1]))

curve(sapply(x, function(phi) ou_nlp_r(phi)), from = log(.1), to = log(10))

ou_nll_tmb <- ou_obj$nll()
ou_nll_tmb$fn(lp_fit$coef)
ou_nll_r(lp_fit$coef)
```

