---
title: "Simulation of SHO model with sine wave noise"
author: "Feiyu Zhu"
date: 
link-citations: true
output: 
  pdf_document:
    number_sections: true
    fig_caption: true
vignette: >
  %\VignetteIndexEntry{sim_sine_wave}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = '80%', fig.align='center')
suppressMessages(require(realPSD))
suppressMessages(require(optimCheck))
suppressMessages(require(testthat))
```

# Derivation of the discrete Fourier transform of sine wave

Given the sine wave function as follows

$$
g(t) = D \sin(2\pi \xi t + \phi), \quad\quad t = 0,\Delta t,\ldots N\Delta t
$$

where $\Delta t = 1/f_s$ ($f_s$ is the sampling frequency).

Its discrete/finite Fourier transform is

$$
\tilde{g}(j) = \frac{D}{\sqrt N} \sum_{n=0}^{N-1} e^{-2\pi i f_j n} \sin(2\pi \xi n\Delta t + \phi)
$$

where $N$ is the total number of discrete observations and $f_j = \frac j N f_s$ ($f_s$ is the sampling frequency).

Recall $\sin(x) = (e^{ix} - e^{-ix})/ 2i$, then we have

$$
\begin{aligned}
\tilde{g}(j) &= \frac{D}{\sqrt N} \sum_{n=0}^{N-1} e^{-2\pi i f_j n\Delta t} \frac{e^{2\pi i \xi n + \phi i} - e^{-2\pi i \xi n - \phi i}}{2i} \\
&= \frac{D}{2i\sqrt N} \left( e^{\phi i} \sum_{n=0}^{N-1} e^{2\pi i (\xi - f_j) n\Delta t} - e^{-\phi i}\sum_{n=0}^{N-1} e^{-2\pi i (\xi + f_j)n\Delta t}  \right) \\ 
&= \frac{D}{2i\sqrt N} \left( e^{\phi i} \frac{e^{2\pi i (\xi - f_j)\Delta t N } - 1}{e^{2\pi i (\xi - f_j)\Delta t} - 1} - e^{-\phi i}\frac{e^{-2\pi i (\xi + f_j)N \Delta t} - 1}{e^{-2\pi i (\xi + f_j)\Delta t} - 1} \right)
\end{aligned}
$$

```{r, sim_sine}
# FFT (discrete Fourier transform) of sine wave noise
fft_sin <- function(fseq, f0, Q, fs, unit_conversion) {
  N <- length(fseq)
  if(!unit_conversion) {
    D <- (Q^0.5) * 3.5e3 / Const
  } else {
    D <- (Q^0.5) * 3.5e3 # 3.5e3 is copied from the MATLAB code
  }
  # k  <- 0.172                  # Cantilever stiffness, N/m
  # Kb <- 1.381e-23              # Boltzmann's constant, (m2*kg)/(s2*K)
  # Temp <- 298                  # Temperature, K
  # D <- 10*psdSHO(f0, f0, Q, k, Kb, Temp, unit_conversion)
  dT <- 1/fs
  xi <- rnorm(1, f0, 10)
  phi <- runif(1, 0, 2*pi)
  sin_fft <- D/(2*1i * sqrt(N)) * (
    exp(phi*1i) * (exp(2*pi*1i*(xi-fseq)*dT*N)-1)/(exp(2*pi*1i*(xi-fseq)*dT)-1) -
    exp(-phi*1i) * (exp(-2*pi*1i*(xi+fseq)*dT*N)-1)/(exp(-2*pi*1i*(xi+fseq)*dT)-1)
  )
  return(sin_fft)
}
```

We can test our derived formula against `R` built-in function `fft` which calcualtes the Discrete Fourier Transform (DFT) of an array by using fast Fourier transform. In R, the DFT is defined as

$$
\tilde{g}(j) = \sum_{n=0}^{N-1} e^{-2\pi i f_j n} D \cdot\sin(2\pi \xi n\Delta t + \phi)
$$

```{r, test_formula}
# testthat tests of the DFT formula
context("Test DFT of Sine Wave")
test_that("Test the formula of DFT of Sine Wave is Correct", {
  ntest <- 20
  for(ii in 1:ntest) {
    N <- sample(10:10000, size = 1)
    N <- 10
    f0 <- 33553
    fs <- 1e7
    dT <- 1/fs
    xi <- rnorm(1, f0, 10)
    phi <- runif(1, 0, 2*pi)
    D <- 3.5e3
    fseq <- (1:N-1)/N * fs
    # method 1: FFT of sin(t)
    sin_time <- D*sin(2*pi * xi * (1:N-1) * dT + phi)
    ans <- fft(sin_time)
    # method 2: formula
    sin_fft <- D/(2*1i*sqrt(N)) * (
      exp(phi*1i) * (exp(2*pi*1i*(xi-fseq)*dT*N)-1)/(exp(2*pi*1i*(xi-fseq)*dT)-1) -
      exp(-phi*1i) * (exp(-2*pi*1i*(xi+fseq)*dT*N)-1)/(exp(-2*pi*1i*(xi+fseq)*dT)-1)
    )
    expect_equal(ans/N, sin_fft/sqrt(N), tolerance = 1e-5)
  }
})
```

# Simulation of sine wave noise directly from frequency domain

First we set up the model constants.

```{r, constants}
# ---------- SHO model parameters ----------
Time  <- 5                   # Total time, second
fs <- 1e6                    # Sampling frequency, Hz
f0 <- 33553                  # Resonance frequency, Hz
Q_vec <- c(1, 10, 100, 500)  # Quality factors, unitless
k  <- 0.172                  # Cantilever stiffness, N/m
Kb <- 1.381e-23              # Boltzmann's constant, (m2*kg)/(s2*K)
Temp <- 298                  # Temperature, K
Const <- 1e30                # unit conversion, 1 m2 = 1e30 fm2
unit_conversion <- TRUE
Aw <- 19000                 # white noise, fm2/Hz 
if(!unit_conversion) Aw <- Aw / Const # if FALSE, we use the standard unit m2/Hz
```

Then we generate the frequency domain.

```{r, freq_domain}
f_lb <- f0 - f0/sqrt(2) # frequency lower bound
f_ub <- f0 + f0/sqrt(2) # frequency upper bound
fseq <- seq(from = f_lb, to = f_ub, by = 1/(10*Time)) # frequency domain, Hz
N <- length(fseq)
```

# Simulation of the SHO periodogram with sine wave noise

We can generate the SHO periodogram from a two dimensional complex multivariate normal distribution. (Since the sum of two squared independent normal with mean 0 and variance $1/2$ is an exponential random variable.) 

```{r, psd_fig_cap, echo = FALSE}
caption <- rep(NA, length(Q_vec))
for(ii in 1:length(Q_vec)) {
  caption[ii] <- paste0("Empirical smoothed periodogram v.s. analytic PSD of SHOW (Q = ", Q_vec[ii], ").")
}
```

We can then plot the smoothed periodogram after binning. (We have converted the unit of PSD from $m^2/Hz$ to $fm^2/Hz$.)

```{r, plot, fig.cap = caption}
bin_size <- 100
for(ii in 1:length(Q_vec)) {
  Q <- Q_vec[ii]
  # psd
  psd <- psdSHO(fseq, f0, Q, k, Kb, Temp, unit_conversion) + Aw
  # sine wave noise
  sin_fft <- fft_sin(fseq, f0, Q, fs, unit_conversion)
  # complex normals
  x1 <- rnorm(N, 0, sqrt(1/2)) * sqrt(psd * fs)
  x2 <- rnorm(N, 0, sqrt(1/2)) * sqrt(psd * fs)
  Y <- complex(real = x1, imaginary = x2)
  # add FFT(sine) directly onto the complex normals
  Y <- (Y + sin_fft) * Conj(Y + sin_fft)
  # Y <- Y * Conj(Y) # for testing purposes
  Y <- Re(Y)
  # smoothed empirical psd (by binning)
  Ybar <- binning(Y, bin_size = bin_size)
  # frequency domain sequence after binning
  fbar <- binning(fseq, bin_size = bin_size)
  # plot
  plot(x = fbar, y = Ybar/fs, 
      xlab = "frequency (Hz)", ylab = expression(paste("PSD (", fm^2/Hz, ")")),
      type = "l", lwd = 0.1, log = "xy")
  lines(x = fbar, y = psdSHO(fbar, f0, Q, k, Kb, Temp, unit_conversion) + Aw, 
      col = "red")
  legend("topright",
    legend = c("Simulated", "Analytic + Aw"), 
    col = c("black", "red"),
    bty = "n", # no box around the legend
    cex = 0.8, # size of the text,
    lwd = 1
  )
}
```

We can also take a closer look at the simulated periodogram for $32kHz < f < 35kHz$. For example, when $Q = 100$, see Figure \ref{fig:small_freq_range}.

```{r, small_freq_range, fig.cap = "Simulated periodogram with $Q = 100$ and added noise."}
Q <- 100
idx <- which(fseq >= f0-1500 & fseq <= f0+1500)
N <- length(fseq)
psd <- psdSHO(fseq, f0, Q, k, Kb, Temp, unit_conversion)
sin_fft <- fft_sin(fseq, f0, Q, fs, unit_conversion)
x1 <- rnorm(N, 0, sqrt(1/2))
x2 <- rnorm(N, 0, sqrt(1/2))
sim_cnorm <- complex(real = x1, imaginary = x2)
Y <- sim_cnorm * sqrt(psd*fs)
Y <- (Y + sin_fft) * Conj(Y + sin_fft)
Y <- Re(Y)
plot(x = fseq[idx]/10^3, y = Y[idx]/fs, xlab = "frequency (kHz)", ylab = expression(paste("PSD (", fm^2/Hz, ")")),
      type = "l", lwd = 0.1, log = "xy")
lines(x = fseq[idx]/10^3, y = psd[idx], 
      col = "red")
legend("topright",
  legend = c("Simulated", "Analytic"), 
  col = c("black", "red"),
  bty = "n", # no box around the legend
  cex = 0.8, # size of the text,
  lwd = 1)
```
