---
title: "Custom PSD models"
author: "Martin Lysy, Feiyu Zhu"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: yes
vignette: >
  %\VignetteIndexEntry{Custom PSD models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<script type="text/x-mathjax-config">
MathJax.Hub.Config({
  TeX: { equationNumbers: { autoNumber: "AMS" } }
});
</script>

\newcommand{\bm}[1]{\boldsymbol{#1}}
\newcommand{\tx}[1]{\textrm{#1}}
\newcommand{\XX}{{\bm{X}}}
\newcommand{\YY}{{\bm{Y}}}
\newcommand{\ZZ}{{\bm{Z}}}
\newcommand{\tth}{{\bm{\theta}}}
\newcommand{\pph}{{\bm{\varphi}}}
\newcommand{\aal}{{\bm{\alpha}}}
\newcommand{\bbe}{{\bm{\beta}}}
\newcommand{\s}{\sigma}
\newcommand{\f}{f}
\newcommand{\ud}{\mathrm{d}}
\newcommand{\psd}{\operatorname{\mathcal S}}
\newcommand{\cov}{\operatorname{cov}}
\newcommand{\var}{\operatorname{var}}
\newcommand{\acf}{\operatorname{\gamma}}
\newcommand{\carfima}{\operatorname{CARFIMA}}
\newcommand{\rv}[3]{#2_{#1},\ldots,#2_{#3}}
\newcommand{\fs}{\f_s}
\newcommand{\kbt}{k_BT}
\newcommand{\Aw}{A_{\tx{w}}}
\newcommand{\Rw}{R_{\tx{w}}}

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
# link to packages
pkg_link <- function(pkg, link) {
  if(link == "github") {
    link <- paste0("https://github.com/mlysy/", pkg)
  } else if(link == "cran") {
    link <- paste0("https://CRAN.R-project.org/package=", pkg)
  }
  paste0("[**", pkg, "**](", link, ")")
}
cran_link <- function(pkg) pkg_link(pkg, "cran")
github_link <- function(pkg) pkg_link(pkg, "github")
```

## Write your own model in C++ using **realPSD**

**realPSD** allows users to easily implement their own model in C++ using **TMB** interface to R. **TMB** (Template Model Builder) is an R package for fitting statistical latent variable models to data, which is extensively documented [here](https://github.com/kaskr/adcomp/wiki/Documentation). **TMB** uses the C++ library for linear algebra [**Eigen**](http://eigen.tuxfamily.org/index.php?title=Main_Page) which is extensively documented [here](http://eigen.tuxfamily.org/dox/). We should be familiar with the basic data types provided by **Eigen** and an entry-level understanding of how to use **TMB** in R.

For any parametric PSD of the form 
$$
\psd(\f,\tth) = \s^2 \cdot U(\f, \pph),
$$
where $\tth = (\s, \pph)$ are the parameters and $\f$ represents a sequence of observations (frequency points), it will be easy for users to define their own model class definition and supply it to TMB for later model fitting.

We will explain the general paradigm through an example.

## An example: Fractional Ornstein-Uhlenbeck Model (compiled on-the-fly)

Here we implement a fractional OU model to make the above steps more concrete. Recall that the fOU model is expressed as
$$
\ud X(t) = -\gamma X(t)\, \ud t + \tau \, \ud B_t^H,
$$
where $\gamma, \tau > 0$ and $B_t^H$ is fractional Brownian motion with Hurst parameter $0 < H < 1$.  The PSD for this model is
\begin{equation}
\label{eq:fcar}
\psd(\f, \tth) = \sigma^2 \times \frac{|\f|^{1-2H}}{\f^2 + \gamma^2},
\end{equation}
where $\pph = (H, \gamma)$ and $\sigma^2 = \tau^2 \cdot \Gamma(2H+1)\sin(\pi H)/(2\pi)$.

The first step to realize the above model is to create a C++ class (`my_model`) which defines two types of members:

1.  Generic public members which have the same name and signature for each user defined class. 
2.  A class constructor which has model-specific input parameters.

The basic structure is given as

```{Rcpp, class, eval = FALSE}
/// @file my_model.hpp
///
/// @brief ...
template <class Type>
  class model_class {
  private:
    //typedefs
    /// Typedef equivalent to `matrix<Type>`.
    typedef Eigen::Matrix<Type, Eigen::Dynamic, Eigen::Dynamic> MatrixXd_t;
    /// Typedef equivalent to `Ref <matrix<Type> >`.
    typedef Eigen::Ref <MatrixXd_t> RefMatrix_t;
    /// Typedef equivalent to `const Ref <const matrix<Type> >`.
    typedef const Eigen::Ref <const MatrixXd_t> cRefMatrix_t;
    // internal variables
    int N_; ///> problem dimensions
    matrix<Type> f_; ///>column vector of frequencies
    matrix<Type> f2_; ///> column vector of squared frequencies.
  public:
    /// Constructor.
    model_class(int N);
    /// Set frequency vector.
    void set_f(cRefMatrix_t& f);
    /// Evaluate the normalized PSD.
    void eval(RefMatrix_t U, cRefMatrix_t& phi);
  };

/// definitions given below
...
...
```

Here let us name `my_model` as `fou`. We may create a C++ header file `fou.hpp` and define the class name as `UFun` (i.e. U function $U(\f, \pph)$). To avoid potential scope conflict, we enclose the class definition within a user-defined namespace `fou`. Thus `model_class` is given by `fou::UFun`. As we can see, some class members are mandatory, e.g. `set_f` and `eval` since they are required in our spectral estimation methods. Users may use different argument names and give them specific definitions based on the specific problems at hand, but these member functions should be defined in the same way as shown above. The private members are defined accordingly (as auxiliary components). For our fOU model, we give the definitions of `set_f(cRefMatrix_t& f)` and `eval(RefMatrix_t U, cRefMatrix_t& phi)` as a reference.

```{Rcpp, fou_class_required_parts, eval = FALSE}
template<class Type>
inline void UFun<Type>::set_f(cRefMatrix_t& f) {
  N_ = f.size();
  f_ = zero_matrix<Type>(N_,1);
  f_ = f;
  f2_ = zero_matrix<Type>(N_,1);
  f2_ = f.cwiseProduct(f);
  return;
}

/// Parameters are: `phi = (H, gamma)`
/// psd = |f|^(1-2H) / (f^2 + gamma^2)
template <class Type>
inline void UFun<Type>::eval(RefMatrix_t U, cRefMatrix_t& phi) {
  U = f2_.array() + phi(1,0) * phi(1,0);
  U = (f_.array().abs()).pow(1-Type(2.0)*phi(0,0)) / U.array(); // use ArrayBase::pow() and ::abs() functions
  return;
}
```

As for the class constructor, since **TMB** macros `DATA_MATRIX()`, etc. don't work inside class member functions, to allow users to decide and change input arguments to the model, we need to define an external constructor `model_ctor(int N, objective_function<Type>* obj)`:

```{Rcpp, external_ctor, eval = FALSE}
#undef TMB_OBJECTIVE_PTR
#define TMB_OBJECTIVE_PTR obj

template<class Type>
UFun<Type> model_ctor(objective_function<Type>* obj) {
  DATA_VECTOR(arg1);
  DATA_SCALAR(arg2); // etc
  return model_class<Type>(N, arg1, arg2);
}

#undef TMB_OBJECTIVE_PTR
#define TMB_OBJECTIVE_PTR this
```

In this way, users may supply their own input arguments with freedom. As developers, we do not need to know these input variables beforehand.

Here we call `model_ctor` as `make_Ufun`. For our fOU model, the definitions of the default class constructor and the external constructor are given below.

```{Rcpp, fou_class_ctor, eval = FALSE}
template<class Type>
inline UFun<Type>::UFun(int N) {
  N_ = N;
  f_ = zero_matrix<Type>(N_, 1);
  f2_ = zero_matrix<Type>(N_,1);
}

#undef TMB_OBJECTIVE_PTR
#define TMB_OBJECTIVE_PTR obj

template<class Type>
UFun<Type> make_Ufun(int N, objective_function<Type>* obj) {
  return UFun<Type>(N);
}

#undef TMB_OBJECTIVE_PTR
#define TMB_OBJECTIVE_PTR this
```

The full souce code of `fou.hpp` is attached at the very end for a complete reference. To warp up the whole process, users only need to do the following things:

1. Come up with a model name `my_model` and create a C++ header file `my_model.hpp` (any helper functions can be included in the same file).
2. Define its `model_class` (possibly within a name scope) by including the default constructor `model_class`, some necessary member functions `set_f` and `eval` and the required private members accordingly.
3. Add an external model class constructor `model_ctor` as previously described.

Now we are ready to compile our template class `fou.hpp` by using `make_psd_model` which automatically supplies the user-defined model to TMB in an R session on-the-fly. Here is the example.

```{r, fou_test_setup, eval = FALSE}
# test generic fou model header created by make_psd_model
require(realPSD)
require(whisker)
require(TMB)

# standalone == TRUE, compile on-the-fly mode
make_psd_model(
  name = "fou", # my_model
  header = "fou.hpp", # name of the header file with extensions, usually it is my_model.hpp
  class = "fou::UFun", # model_class with name scope (if any)
  ctor = "fou::make_Ufun", # model_ctro with name scope (if any)
  include = "realPSD/FitMethods.hpp", # to use the pre-compiled methods provided by realPSD
  method = "realPSD::FitMethods", # to use the pre-compiled methods provided by realPSD
  standalone = TRUE,
  path = "fou_Generics.cpp"
)

compile("fou_Generics.cpp", PKG_CXXFLAGS = paste0("-I", system.file("include", package = "realPSD")))
dyn.load(dynlib("fou_Generics"))
model <- "fou_Generics"
```

Now we explain the above input arguments of `make_psd_model` in detail:

- `name`: Name of `my_model`.
- `header`: Name of the header file with extensions, usually it is `my_model.hpp`.
- `class`: `model_class` with name scope (if any).
- `ctor`: `model_ctor` with name scope (if any).
- `include`: Additional include directive path, usually package specific, in this example, the path to `FitMethods.hpp`.
- `method`: Generic method class with enclosing namespace if it exists, to use our package realPSD, then it's always realPSD::FitMethods.
- `standalone`: If `TRUE` creates a standalone `cpp` file to pass to [TMB::compile()].  Otherwise, creates an `hpp` header file to be placed in a package created with **TMBtools**.
- `path`: Name of output file includeing path.

We can test the model by using the following unit tests:

```{r, fou_test_on_the_fly, eval = FALSE}
require(testthat)
# helper function: calculate the psd in R
fou_psd_r <- function(f, phi) {
 psd <- abs(f)^(1-2*phi[1]) / (f^2 + phi[2]^2)
}

context("Test: fractional OU model")

test_that("The UFun returned by TMB (C++) is the same as that given by R", {
  ntests <- 20
  for(ii in 1:ntests) {
    N <- sample(10:20, 1)
    f <- matrix(runif(N, 0, 2*N))
    phi <- matrix(c(runif(1), rexp(1))) # phi = c(H, gamma) where 0 < H < 1, gamma > 0
    obj <- MakeADFun(data = list(f = f, method = "UFun"),
                 parameters = list(phi = matrix(c(runif(1), rexp(1)))),
                 DLL = model, silent = TRUE)
    psd_tmb <- obj$simulate(c(phi))$U
    psd_r <- fou_psd_r(f, phi)
    expect_equal(psd_tmb, psd_r)
  }
})

# helpfer function
mle_tau_r <- function(f, Y, phi, ufun, fs) {
  U <- fs * ufun(f, phi)
  mean(Y/U)
}
sim_Y <- function(n) rchisq(n, df = 2)
# normalized PSD
# phi = c(H, gamma)
fou_ufun <- function(f, phi) {
  abs(f)^(1-2*phi[1]) / (f^2 + phi[2]^2)
}
# simulate the sampling freq
sim_fs <- function() sample(10:1000, size = 1)

test_that("sigma^2 returned by TMB (C++) is the same as that given by R", {
  ntests <- 20
  for(ii in 1:ntests) {
    N <- sample(50:100, 1)
    f <- matrix(runif(N, 0, 2*N))
    fs <- sim_fs()
    phi <- matrix(c(runif(1), rexp(1))) 
    Y <- matrix(sim_Y(N))
    # TMB obj
    obj <- MakeADFun(
      data = list(f = f, Y = Y, fs = fs, method = "MLE_tau"), 
      parameters = list(phi = matrix(0.5, 2)), 
      DLL = model, silent = TRUE
    )
    # define mle_tau from TMB
    mle_tau_tmb <- function(phi) obj$fn(phi)
    # comparison
    tau_tmb <- mle_tau_tmb(phi)
    tau_r <- mle_tau_r(f, Y, phi, ufun = fou_ufun, fs)
    expect_equal(tau_tmb, tau_r)
  }
})
```
Please feel free to try to design your own model and fit its parameters by using the three different methods (MLE, NLS, LP) provided by **realPSD**. Have fun!

## Write your own package

If you are trying to include such a TMB model in your own R package and to make use of the `realPSD` methods, it is easy to do with `make_psd_model` by only changing `standalone` to `FALSE` and then apply `TMBtools::export_models()`. For more details, please see [here](https://github.com/mlysy/TMBtools.git).

## Summary

`realPSD` provides three different spectral estimation methods as well as an automatic mechanism to help users supply their own model to TMB either on-the-fly or in their own package. The general design paradigm of such a mechanism can be summarized as follows:

- User writes `my_model` class which defines two types of public members:

  1.  Generic public members which have the same name and signature for each user defined class `my_model_A`, `my_model_B`, etc.

  2.  A class constructor which has model-specific input parameters.

- Our package `generic_pkg` contains generic code which can operate on an instantiated object `my_model` of type `my_model<Type>`.

- Package automatically creates a **TMB** file out of user and our package code, i.e., which can be compiled on-the-fly or added to the user's R/**TMB** package created with **TMBtools**.

The idea of defining an external class constructor is the key to give users the freedom to add their own input arguments, which overcomes the limitation of **TMB** in R. The generic code relies upon `whisker`, a {{Mustache}} implementation in R to realize logic-less templates.

## Appendix

Here is the source code of the user-defined model header file `fou.hpp`.

```{Rcpp, fou, eval = FALSE}
/// @file fou.cpp
///
/// @brief fractional OU model, passing data to TMB
namespace fou {
  /// create `matrix<Type>` of zeros as a helper function
  ///
  /// @param[in] n Integer number of rows.
  /// @param[in] p Integer number of columns.
  /// @return A `matrix<Type>` of size `n x p` initialized with zeros.
  template<class Type>
  matrix<Type> zero_matrix(int n, int p) {
    matrix<Type> out(n,p);
    out.setZero();
    return out;
  }

  /// A template class for fractional OU model
  template <class Type>
  class UFun {
  private:
    //typedefs
    /// Typedef equivalent to `matrix<Type>`.
    typedef Eigen::Matrix<Type, Eigen::Dynamic, Eigen::Dynamic> MatrixXd_t;
    /// Typedef equivalent to `Ref <matrix<Type> >`.
    typedef Eigen::Ref <MatrixXd_t> RefMatrix_t;
    /// Typedef equivalent to `const Ref <const matrix<Type> >`.
    typedef const Eigen::Ref <const MatrixXd_t> cRefMatrix_t;
    // internal variables
    int N_; ///> problem dimensions
    matrix<Type> f_; ///>column vector of frequencies
    matrix<Type> f2_; ///> column vector of squared frequencies.
  public:
    /// Constructor.
    UFun(int N);
    /// Set frequency vector.
    void set_f(cRefMatrix_t& f);
    /// Evaluate the normalized PSD.
    void eval(RefMatrix_t U, cRefMatrix_t& phi);
  };

  template<class Type>
  inline UFun<Type>::UFun(int N) {
    N_ = N;
    f_ = zero_matrix<Type>(N_, 1);
    f2_ = zero_matrix<Type>(N_,1);
  }

  template<class Type>
  inline void UFun<Type>::set_f(cRefMatrix_t& f) {
    N_ = f.size();
    f_ = zero_matrix<Type>(N_,1);
    f_ = f;
    f2_ = zero_matrix<Type>(N_,1);
    f2_ = f.cwiseProduct(f);
    return;
  }

  /// Parameters are: `phi = (H, gamma)`
  // psd = |f|^(1-2H) / (f^2 + gamma^2)
  template <class Type>
  inline void UFun<Type>::eval(RefMatrix_t U, cRefMatrix_t& phi) {
    U = f2_.array() + phi(1,0) * phi(1,0);
    U = (f_.array().abs()).pow(1-Type(2.0)*phi(0,0)) / U.array(); // use ArrayBase::pow() and ::abs() functions
    return;
  }

  #undef TMB_OBJECTIVE_PTR
  #define TMB_OBJECTIVE_PTR obj

  template<class Type>
  UFun<Type> make_Ufun(int N, objective_function<Type>* obj) {
    return UFun<Type>(N);
  }

  #undef TMB_OBJECTIVE_PTR
  #define TMB_OBJECTIVE_PTR this
} // end namespace fou
```

## Create the OU model object in C++

To provide efficient estimation for general parametric PSD models, **realPSD** requires you to write your own model in a standalone C++ header file (with extension `.h` or `.hpp`). Here we have it ready for you. 

```{r ou_ufun, echo = FALSE, results = "asis"}
cat("```cpp", readLines("OU_Model.hpp"), "```", sep = "\n")
```

Basically, we defined a `UFun` class within our own namespace `ou` to avoid any potential scope conflict with the **realPSD** package. Later whenever we refer to this class, we will use `ou::UFun` to emphasize its namespace. The `ou::UFun` class is nothing fancy but just a collection of necessary data and functions we need to evaluate the likelihood function of the OU process. In our specific example, we need to provide the length of the simulated time series `N` (which is supplied via the constructor function `UFun(int N)`), a vector of frequencies `f` (which is supplied via function `set_f`) and a function `eval` to evaluate the PSD. If you look at the definition of `eval` function

```C++
/// @param[out] U Normalized PSD vector of size `N`.
/// @param[in] phi Computational basis parameters.  Here we have `phi = log(alpha)`.
template <class Type>
inline void UFun<Type>::eval(RefMatrix<Type> U, cRefMatrix<Type>& phi) {
  U = (scale_ * f2_).array() + exp(Type(2.0)*phi(0,0));
  // U = exp(Type(2.0)*phi(1,0)) / U.array();
  U = U.array().inverse();
  return;
}
```
you will notice it is basically the C++ translation of the R function `ou_psd`. The only difference is that for numerical purposes, we define the PSD as $S(f) = \beta^2 U(f)$ where 
$$
U(f) = \frac{1}{(2\pi f)^2 + \alpha^2}
$$
is the normalized PSD we are using in **realPSD**.

The private member variables `f_`, `f2_` and `scale_` can be understood as some intermediary variables to help us create the function `eval` eventually. But we don't need them to be visible outside of the class `ou::UFun`, so they are private. These variables are by no means mandatory but just for convenience. You should think about how to create your own private members based on your specific model. 

As for `RefMatrix<Type>`, it is a user-defined data type used by **realPSD** to help users make use of data types in C++ [Eigen](http://eigen.tuxfamily.org/index.php?title=Main_Page) library. You can understand `RefMatrix<Type>` simply as a built-in data type for matrix. Please make sure you have the line `#include "realPSD/utils.hpp"` at the beginning of your header file. Then you header file will automatically include some tools to facilitate your job.

You may also notice that we have an external constructor `template<class Type> UFun<Type> make_Ufun(...)` for `ou::UFun` class. Just copy and paste this function to the end of your header file (still within the namespace), it will help you get around some technical issues involved in `r github_link("TMB")`. For more details, please see the [advanced tutorial](http://htmlpreview.github.com/?https://github.com/mlysy/realPSD/master/doc/realPSD-advancedtut.html).

To summarize, you can follow the steps below to create a model header file:

1. Create an empty C++ header file with extension `.h` or `.hpp` and name it whatever you like. In our example, we name the file `OU_Model.hpp`. Put it under the current working directory.
2. Give a name to your namespace to wrap up your model class. Usually the namespace is named according to your model, e.g. `ou` in our example.
3. Create a `UFun` class within your namespace. It is recommended to define the class as a template class in the way as shown in our example, i.e. `template <class Type>
  class UFun {...};`. It would enable you to make use of the data types given by C++ [Eigen](http://eigen.tuxfamily.org/index.php?title=Main_Page) library. 
4. Create necessary public members (variables or functions) to get data from outside and to evaluate your normalized PSD, i.e. the U(f) function and necessary private members for internal computation. Usually it is a good idea to make your own constructor (instead of using the default constructor) to initialize some private data members as we did in the example. 
5. The rest of the file should be just a simple copy-and-paste of the above example header file.

